---
title: "Welcome"
editor: visual
---

This site contains R code and some explanatory text to help with completing the data analyses for ANIM5502 in 2022.

**Important!**

ANIM5502 is not a coding nor a data analysis course - you have the *Big Data* course for those!

The purpose of this website is to provide all the code necessary for running your analyses for the ANIM5502 group project. If you find useful coding tips here then that is great, however, the main thing is to see how we use code/data analyses to answer **interesting biological questions**.

So explore the site, use it to help with your data analyses, but don't get bogged down in the code!

## How to use this site

Use the menu on the left to navigate the site.

Code from each section can be copied using the clipboard symbol on the right of each 'code block' (the grey boxes). Function help pages can be accessed by clicking on a function name. You can try this below.

```{r}
#| output: false
print("Success - too easy to copy and paste code")
```

## How to download and run the R scripts on your computer

If you are not a git/github user, you can use the following instructions to download and run the code on your computer.:

### First create an RStudio project

Open RStudio and then follow these steps:

-   Select the **File** menu followed by **New Project**

-   Choose **New Directory** and then **New Project**

-   Give the directory a name, for example, **ANIM5502_2022_local**, choose where to save the folder by selecting **Browse**, check the box next to **Open in new session**. Select **Create Project**.

### Then run the following code whilst in the opened project

**Warning**

Running the following code will download all the R scripts as well as the prepared data. Do not run the code if you don't want this to happen!

Note: it will not download the Excel file that is needed to run the data_cleaning code.

```{r}
#| eval: false

download.file("https://raw.githubusercontent.com/jdberson/ANIM5502_2022/main/code/ANIM_5502_local_setup.R",
              "ANIM_5502_local_setup.R")

source("ANIM_5502_local_setup.R")

```
